### 2022-11-23 response-letter.md
Responses 1 & 5 [Review_02] response letter #89

Dear editor,

Thank you for the inviting a specialist on large scale biodiversity patterns of
Africa to review our manuscript. We were happy to see that the first reveiwer
was satisfied with our response, and that the second reviewer also saw value in
our contribution. The new comments were very pertinent and we hope that this
letter and the changes we've made to the paper are satisfactory.

We have improved the text in order to avoid misleading phrasing. For example, we
understand that using the words "updated ranges" can be read as distribution
modelling, which is not the case: here, we simply use a method to identify areas
where the predator range does not match with their prey's range, which might
indicate distribution uncertainty areas.

We have also addressed reviewer's concerns about the GBIF data (see below): our investigations show that most of the data we were using are recent (from the last 20 years). We describe the time distribution of GBIF records to explicit that. Finally, we have improved the methods and the discussion sections to include more detailed descriptions of the data being used and to give more context to the results. We detail specific changes below (line numbers refer to the clean, revised PDF file).

>*"1. In the extent of the study, the authors mention that they “restricted the rasters a spatial extent comprised between latitudes 35°S and 40°N and longitudes 20°W and 55°E. “ – which is the whole of Africa. Since the study uses a dataset from Tanzania (Serengeti food web dataset), roughly 3% of the continent, how did the authors deal with the presence of putative prey items in other parts of the continent that are absent from the Serengeti food web dataset? For example Papio ursinus, Sylvicapra caffra and S. grimmia don't occur in Tanzania but are prey items to many carnivores in Africa."*

>*"5. Taxonomic groups: Carnivores feed on a diversity of items, beyond herbivores. Their diet might include birds, fish, reptiles, amphibians, and even invertebrates. How did the authors account for the fact that despite missing the herbivores from Serengeti, there were other sources of protein that could sustain the species in the continent?"*

We have found that these two comments address the same issue: the representativity of the Serengeti foodweb. First, we would like to emphasize that our analysis is meant to be a proof of concept, showing the consequences of data limitation. Therefore, rather than trying to gather an exhaustive foodweb dataset - with information for all possible interactions for all occurring mammals in the African continent -, we looked for a very complete, readily available, widely used and well known dataset which could represent the main interactions in that area. Interaction data is difficult to collect, especially at broad scales. Because of that, we expected that no foodweb would be complete, and we are aware of the biases of our interactions data. However, the goal of this paper is to investigate precisely what happens if we need to deal with incomplete and biased datasets, because we believe that this might lead us to describe unrealistic ecological communities at the local scale. More than that, we believe that data incompleteness is not a limitation of our study, but one of its strengths, because we present a simple method to identify and describe these limitations.

That being said, the foodweb database that we chose is known for being representative for these animals on the whole African continent (as we can see, for example, in de Visser, Freymann and Olff, 2011). Although our dataset miss a few links, again, it was not our goal to make a comprehensive analysis of all the area where a predator would miss a prey based on all the knowledge available or not, but rather demonstrate what can happen if we deal with incomplete datasets, highlight the importance of accounting for uncertainty when dealing with species distribution data, and call to attention the need to make more interaction data open. Moreover, many of the prey items on Baskerville's dataset were also found outside the Serengeti park area, and our method only needed one link connecting a predator to a herbivore to "keep" a pixel from a predator's range. Therefore, the impact of the missing links on the foodweb should be of great magnitude to invalidate our methods.

Finally, some of the results we highlight would not change with the addition of new prey items. For example, the *Canis aureus* results are due mainly to an apparent taxonomic disagreement - which is indeed connected to its interaction data, but this would not change with a more complete list of preys for this species.

To emphasize and clarify these points, we have changed lines XX, XX and XX on the manuscript.

We hope our revised manuscript is now more appropriate for publishing at PeerJ.
Reiterating our appreciation for yours and the reviewer's thoughtful comments,
we are available to answer any further questions.

Best regards,
The Authors