### 2022-11-23 response-letter.md
Forgot to wrap, sorry!

Dear editor,

Thank you for the inviting a specialist on large scale biodiversity patterns of
Africa to review our manuscript. We were happy to see that the first reveiwer
was satisfied with our response, and that the second reviewer also saw value in
our contribution. The new comments were very pertinent and we hope that this
letter and the changes we've made to the paper are satisfactory.

We have improved the text in order to avoid misleading phrasing. For example, we
understand that using the words "updated ranges" can be read as distribution
modelling, which is not the case: here, we simply use a method to identify areas
where the predator range does not match with their prey's range, which might
indicate distribution uncertainty areas.

We have also addressed reviewer's concerns about the GBIF data (see below): our
investigations show that most of the data we were using are recent (from the
last 20 years). We describe the time distribution of GBIF records to explicit
that. Finally, we have improved the methods and the discussion sections to
include more detailed descriptions of the data being used and to give more
context to the results. We detail specific changes below (line numbers refer to
the clean, revised PDF file).

>*"1. In the extent of the study, the authors mention that they “restricted the
>rasters a spatial extent comprised between latitudes 35°S and 40°N and
>longitudes 20°W and 55°E. “ – which is the whole of Africa. Since the study
>uses a dataset from Tanzania (Serengeti food web dataset), roughly 3% of the
>continent, how did the authors deal with the presence of putative prey items in
>other parts of the continent that are absent from the Serengeti food web
>dataset? For example Papio ursinus, Sylvicapra caffra and S. grimmia don't
>occur in Tanzania but are prey items to many carnivores in Africa."*

>*"5. Taxonomic groups: Carnivores feed on a diversity of items, beyond
>herbivores. Their diet might include birds, fish, reptiles, amphibians, and
>even invertebrates. How did the authors account for the fact that despite
>missing the herbivores from Serengeti, there were other sources of protein that
>could sustain the species in the continent?"*

We have found that these two comments address the same issue: the
representativity of the Serengeti foodweb. First, we would like to emphasize
that our analysis is meant to be a proof of concept, showing the consequences of
data limitation. Therefore, rather than trying to gather an exhaustive foodweb
dataset - with information for all possible interactions for all occurring
mammals in the African continent -, we looked for a very complete, readily
available, widely used and well known dataset which could represent the main
interactions in that area. Interaction data is difficult to collect, especially
at broad scales. Because of that, we expected that no foodweb would be complete,
and we are aware of the biases of our interactions data. However, the goal of
this paper is to investigate precisely what happens if we need to deal with
incomplete and biased datasets, because we believe that this might lead us to
describe unrealistic ecological communities at the local scale. More than that,
we believe that data incompleteness is not a limitation of our study, but one of
its strengths, because we present a simple method to identify and describe these
limitations.

That being said, the foodweb database that we chose is known for being
representative for these animals on the whole African continent (as we can see,
for example, in de Visser, Freymann and Olff, 2011). Although our dataset miss a
few links, again, it was not our goal to make a comprehensive analysis of all
the area where a predator would miss a prey based on all the knowledge available
or not, but rather demonstrate what can happen if we deal with incomplete
datasets, highlight the importance of accounting for uncertainty when dealing
with species distribution data, and call to attention the need to make more
interaction data open. Moreover, many of the prey items on Baskerville's dataset
were also found outside the Serengeti park area, and our method only needed one
link connecting a predator to a herbivore to "keep" a pixel from a predator's
range. Therefore, the impact of the missing links on the foodweb should be of
great magnitude to invalidate our methods.

Finally, some of the results we highlight would not change with the addition of
new prey items. For example, the *Canis aureus* results are due mainly to an
apparent taxonomic disagreement - which is indeed connected to its interaction
data, but this would not change with a more complete list of preys for this
species.

To emphasize and clarify these points, we have changed lines XX, XX and XX on
the manuscript.

We hope our revised manuscript is now more appropriate for publishing at PeerJ.
Reiterating our appreciation for yours and the reviewer's thoughtful comments,
we are available to answer any further questions.

Best regards,
The Authors